---
title: "class5_Quarto"
author: "Yunchong Zhao"
format: pdf
toc: TRUE
---

# Our First Plot

R has basic graphics

```{r}
plot(cars)
```

## How would I plot this with `ggplot2`?

```{r}
library("ggplot2")

ggplot(cars)+
  aes(speed, dist)+
  geom_point() +
  #blinker on a particular funciton and press F1 to get to the help page
  geom_smooth(method="lm", se=FALSE) +
  labs ( title = "speed vs. stopping distance",
         x= "speed(MPH)",
         y= "stopping distance(ft)",
         subtitle = "subtitle",
         caption ="   from dataset \"Cars\"   " ) +
  theme_bw()
```

```{r}
#introducing the genes dataset
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)

head(genes)
nrow(genes)
colnames(genes)
ncol(genes)

table(genes$State)

#what is the percentage of genes being upregulated?
round( 127/5196*100, 2)
round(table(genes$State)/nrow(genes)*100, 2)

#ggplot practices for genes
library("ggplot2")
p <- ggplot(genes) +
      aes(Condition1, Condition2, col=State)+
      geom_point()

p_refine <- p + 
            scale_color_manual( values=c("blue", "gray", "red")) +
            labs( x="Control (no drug)", 
                  y="Drug Treatment",
                  title="Gene Expression Changes Upon Drug Treatment",
                  )
p_refine


```

Below is an example of "in-line R-code":\
\
There are `r nrow(genes)` genes in this dataset

# Going Further

```{r}
library("gapminder")
library("ggplot2")
library("dplyr")
# File location online
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"

gapminder <- read.delim(url)

#ggplot for 2007
gapminder_2007 <- gapminder %>% filter(year==2007)

ggplot_2007 <- ggplot(gapminder_2007)+
  aes(gdpPercap, lifeExp, color=continent, size=pop)+
  geom_point(alpha=0.7) +
  scale_size_area(max_size=15)

#print 2007 ggplot
ggplot_2007

#ggplot for 1957
gapminder_1957 <- gapminder %>% filter(year==1957)

ggplot_1957 <- ggplot(gapminder_1957)+
  aes(gdpPercap, lifeExp, color=continent, size=pop)+
  geom_point(alpha=0.7) +
  scale_size_area(max_size=15)

#print 1957 ggplot
ggplot_1957

#bringing two plots together
gapminder_50years <- gapminder %>% filter(year==1957 | year==2007)

ggplot_50years <- ggplot(gapminder_50years)+
  aes(gdpPercap, lifeExp, color=continent, size=pop)+
  geom_point(alpha=0.7) +
  scale_size_area(max_size=15)+
#using the facet funciton 
  facet_wrap(~year)

#print 50year difference
ggplot_50years
```

You can add options to executable code like this

```{r echo=FALSE}
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

#Bar Graphs

```{r}
library("gapminder")
library("ggplot2")

# File location online
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"

gapminder <- read.delim(url)

gapminder_top5 <- gapminder %>% 
  filter(year==2007) %>% 
  arrange(desc(pop)) %>% 
  top_n(5, pop)

gapminder_top5

#bar chart ggplot

barplot_top5_bar <- ggplot(gapminder_top5)+
    geom_col(aes(x= reorder(country, -pop), y=pop, fill= country ), col="gray30")+
  guides(fill="none")


barplot_top5_bar
```

```{r}
#to not run this chunk:
#| eval: false

```

# Flipping Bar Charts

```{r}
library("ggplot2")

head(USArrests)

#flipping bar chart
USArrests$State <- rownames(USArrests)
ggplot(USArrests) +
  aes(x=reorder(State,Murder), y=Murder) +
  geom_col() +
  coord_flip()

# better aesthetics
USArrests$State <- rownames(USArrests)
ggplot(USArrests) +
  aes(x=reorder(State,Murder), y=Murder) +
  geom_point() +
  geom_segment( aes(x=State,
                    xend=State,
                    y=0,
                    yend=Murder), color="blue" )+
  coord_flip()
```

# GGplot Animation

```{r}
#install.packages("gifski")       ##remove # if need installation
#install.packages("gganimate")    ##remove # if need installation
library("gifski")
library("gganimate")

# File location online
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"

gapminder <- read.delim(url)

# Setup nice regular ggplot of the gapminder data
ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  # Facet by continent
  facet_wrap(~continent) +
  # Here comes the gganimate specific bits
  labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
  transition_time(year) +
  shadow_wake(wake_length = 0.1, alpha = FALSE)
```

# Combining Plots

```{r}
#install.packages("patchwork")  ##remove # if need installation

library(patchwork)

# Setup some example plots 
p1 <- ggplot(mtcars) + geom_point(aes(mpg, disp))
p2 <- ggplot(mtcars) + geom_boxplot(aes(gear, disp, group = gear))
p3 <- ggplot(mtcars) + geom_smooth(aes(disp, qsec))
p4 <- ggplot(mtcars) + geom_bar(aes(carb))

# Use patchwork to combine them here:
(p1 | p2 | p3) /
      p4
```

# Session Info

```{r}
sessionInfo()
```
